Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-08-14 Master Thesis/synopsis.tex
Fix
Avoid iff

\section{The Valuation Spectra}

@@

\begin{proof}
By Lemma \ref{2.2-pre}, $w$ induces a valuation $\tilde{w}\in\operatorname{Spv}k(v)$. Let $k(v)_{\tilde{w}}$ be henselization of $k(v)$ with respect to $\tilde{w}$. By Theorem \ref{theorem-absoluteinertia}, $k(v)_{\tilde{w}}$ corresponds to an intermediate extension $K_v\subset L \subset K^\text{t}_v$. Let $v_L$ be the unique extension of $v$ on $L$. Then $k(v_L)=k(v)_{\tilde{w}}$. Let $w_L$ be the valuation on $L$ corresponding to $\tilde{w}\in \operatorname{Spv} k(v_L)$. By \cite{EP05} Corollary 4.1.4, $w_L$ is henselian iff both $v_L$ and $\tilde{w}$ are henselian, thus $L$ is a henselization of $w$.
\end{proof}

@@

\begin{remark}
Two valuations of rank $1$ are independent iff they are inequivalent.
\end{remark}

\begin{corollary}\label{corollary-fkschmidt}
Let $w_1,w_2\in\operatorname{Spv}(K^\text{s})$ be two inequivalent valuations of rank $1$ with decomposition groups $G_{{w_1}},G_{{w_2}}\subset G_K$. Then $G_{{w_1}}\cap G_{{w_2}}=1$.
\end{corollary}

\begin{proof}
Let $L$ be the fixed field of $G_{w_1}\cap G_{w_2}\subset G_K$ and set $v_i = w_i|_L$. Then $L$ is henselian with respect to both $v_1$ and $v_2$. Since $w_1,w_2$ are the only extensions of $v_1$ and $v_2$ in $K^\text{s}$ respectively, $v_1$ and $v_2$ are inequivalent valuations of rank $1$. Now $L$ is separably closed by Theorem \ref{theorem-fkschmidt}. Therefore, by Theorem \ref{theorem-henselization}, we get $G_{{w_1}}\cap G_{{w_1}}=G_L =1$.
\end{proof}

@@

\begin{lemma}\label{2.2}
Let $K$ be henselian with respect to two different valuations $w_1,w_2\in |\operatorname{Spv}(K)|$. Then $K$ is algebraically closed or $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K),\ v(w_1)=v(w_2)$ and $k(v(w_1))$ is separably closed. In this case, we have $\operatorname{cd}(K)\leq 2$.
\end{lemma}

@@

\begin{proof}
Let $L/K$ be a finite subextension of $\overline{K}/K$. Let $\ell$ and $\ell'$ be the constant fields of $L$ and $L'$ respectively. Applying Lemma \ref{4.2-pre} to $K$ and $L$ gives:
\[ [\ell:k] = [\ell':k'] \]
Therefore $L/K$ is a constant field extension iff $L'/K'$ is a constant field extension. Since $K\overline{k}/K$ and $K'\overline{k'}/K$ are the union of their finite subextensions, we obtain $(K\overline{k})' = K'\overline{k'}$. This implies that $\sigma$ induces an isomorphism:
\[G_k \cong \operatorname{Gal}(K\overline{k}/K) \to \operatorname{Gal}(K'\overline{k'}/K')\cong G_{k'}\] 
Everything else follows from Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}

@@

\begin{lemma}\phantomsection\label{5.2}
\begin{enumerate}[(i)]
\item Let $f: A\to B$ be a homomorphism of abelian groups with finite kernel and cokernel. Then $A$ is free-by-finite iff $B$ is free-by-finite.
\item Let $0\longrightarrow A'\stackrel{i}{\longrightarrow}A\stackrel{p}{\longrightarrow}A''\longrightarrow 0$ be an exact sequence of abelian groups. If $A'$ and $A''$ are free-by-finite, then $A$ is free-by-finite. 
\item Let $0\longrightarrow A'\stackrel{i}{\longrightarrow}A\stackrel{p}{\longrightarrow}A''\longrightarrow 0$ be an exact sequence of abelian groups. If $A$ is free-by-finite and $A'$ is finitely generated, then $A''$ is free-by-finite.
\item Let $A$ be free-by-finite. Then $\bigcap_{n\in\mathbb{N}}nA=0$.
\end{enumerate}
\end{lemma}

@@

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ set:
\begin{align*}
K^S &= \bigcap_{P\in X_0 -  S} U_P
\end{align*}
$K'^{S'}$ is defined similarly. For finite subsets $S\subset X_0$ the corresponding subset of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}

>>>>

\section{The Valuation Spectrum}

@@

\begin{proof}
By Lemma \ref{2.2-pre}, $w$ induces a valuation $\tilde{w}\in\operatorname{Spv}k(v)$. Let $k(v)_{\tilde{w}}$ be henselization of $k(v)$ with respect to $\tilde{w}$. By Theorem \ref{theorem-absoluteinertia}, $k(v)_{\tilde{w}}$ corresponds to an intermediate extension $K_v\subset L \subset K^\text{t}_v$. Let $v_L$ be the unique extension of $v$ on $L$. Then $k(v_L)=k(v)_{\tilde{w}}$. Let $w_L$ be the valuation on $L$ corresponding to $\tilde{w}\in \operatorname{Spv} k(v_L)$. By \cite{EP05} Corollary 4.1.4, $w_L$ is henselian if and only if both $v_L$ and $\tilde{w}$ are henselian, thus $L$ is a henselization of $w$.
\end{proof}

@@

\begin{remark}
Two valuations of rank $1$ are independent if and only if they are inequivalent.
\end{remark}

\begin{corollary}\label{corollary-fkschmidt}
Let $v_1,v_2\in\operatorname{Spv}(K^\text{s})$ be two inequivalent valuations of rank $1$ with decomposition groups $G_{{v_1}},G_{{v_2}}\subset G_K$. Then $G_{{v_1}}\cap G_{{v_2}}=1$.
\end{corollary}

\begin{proof}
Let $L$ be the fixed field of $G_{v_1}\cap G_{v_2}\subset G_K$. Then $L$ is henselian with respect to both $v_1|_L$ and $v_2|_L$. Since $v_1,v_2$ are the only extensions of $v_1|_L$ and $v_2|_L$ in $K^\text{s}$ respectively, $v_1$ and $v_2$ are inequivalent valuations of rank $1$ as well. Now $L$ is separably closed by Theorem \ref{theorem-fkschmidt}. Therefore, by Theorem \ref{theorem-henselization}, we get $G_{{v_1}}\cap G_{{v_1}}=G_L =1$.
\end{proof}

@@

\begin{lemma}\label{2.2}
Let $K$ be henselian with respect to two different valuations $w_1,w_2\in |\operatorname{Spv}(K)|$. Then $K$ is algebraically closed or $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K),\ v(w_1)=v(w_2)$, and $k(v(w_1))$ is separably closed. In this case, we have $\operatorname{cd}(K)\leq 2$.
\end{lemma}

@@

\begin{proof}
Let $L/K$ be a finite subextension of $\overline{K}/K$. Let $\ell$ and $\ell'$ be the constant fields of $L$ and $L'$ respectively. Applying Lemma \ref{4.2-pre} to $K$ and $L$ gives:
\[ [\ell:k] = [\ell':k'] \]
Therefore $L/K$ is a constant field extension if and only if $L'/K'$ is a constant field extension. Since $K\overline{k}/K$ and $K'\overline{k'}/K$ are the union of their finite subextensions, we obtain $(K\overline{k})' = K'\overline{k'}$. This implies that $\sigma$ induces an isomorphism:
\[G_k \cong \operatorname{Gal}(K\overline{k}/K) \to \operatorname{Gal}(K'\overline{k'}/K')\cong G_{k'}\] 
Everything else follows from Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}

@@

\begin{lemma}\phantomsection\label{5.2}
\begin{enumerate}[(i)]
\item Let $f: A\to B$ be a homomorphism of abelian groups with finite kernel and cokernel. Then $A$ is free-by-finite if and only if $B$ is free-by-finite.
\item Let $0\longrightarrow A'\stackrel{i}{\longrightarrow}A\stackrel{p}{\longrightarrow}A''\longrightarrow 0$ be an exact sequence of abelian groups. If $A'$ and $A''$ are free-by-finite, then $A$ is free-by-finite. 
\item Let $0\longrightarrow A'\stackrel{i}{\longrightarrow}A\stackrel{p}{\longrightarrow}A''\longrightarrow 0$ be an exact sequence of abelian groups. If $A$ is free-by-finite and $A'$ is finitely generated, then $A''$ is free-by-finite.
\item Let $A$ be free-by-finite. Then $\bigcap_{n\in\mathbb{N}}nA=0$.
\end{enumerate}
\end{lemma}

@@

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ we set:
\begin{align*}
K^S &= \{ x\in K^\times \mid v_P(x)=0\quad\text{for all }P\in X_0-S \}
\end{align*}
$K'^{S'}$ is defined similarly. For finite subsets $S\subset X_0$ the corresponding subset of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}
